Trending assessment for nowcasts, measurements, and forecasts

### 2024-03-04 sections/03-trending.tex
Update on Overleaf.

All measures can also be evaluated as a rolling estimate to detect changes in performance over time.
A rolling estimate is a sequence of estimates, each computed on a length-$w$-window of the data.
For the trending ratio, a rolling estimate with a backward-looking window is given by
\begin{equation*}
    \acc_t (\diffx, \diffy) \coloneqq \frac{\sum_{t^\star = t-w + 1}^{t} \ind{\diffxt[t^\star] \diffyt[t^\star] > 0}}{w} \quad t = w-1, \dots, T.\label{eq:acc_rolling}
\end{equation*}
Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the trending ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and~\ref{fig:trending_basic_4q_sample_color}.
Thus, the yearly course of the prediction's trending ratio can be detected.
The trending ability of the signal has a strong yearly, sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
This seasonal behavior cannot be observed in the two four-quadrant plots as the high number of points overlays it or in the trending ratio.

@@

In applications, signal data is often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-forecasting}).
We refer to time steps, for which either signal or true values or both are unavailable, as missing values.
Data pairs with missing values can be excluded from the computation of the measures so that the time indices with missing values are excluded from $\mathcal{T}$.
In general, the measures can be computed for data with missing values if the missing values do not systematically occur.
Systematic missing values would lead to a biased estimate of the trending ratio, for example, if signal data is omitted in times of high change and thus uncertainty.
In this case, the measures should be computed for the time steps with available data, and the results should be interpreted cautiously.
Thus, inspection of missing data should accompany applications with missing data, for example, through a visual assessment.
In the applications in Sections~\ref{sec:application-covid} and~\ref{sec:application-forecasting}, only a few missing values occur, and we give information on the missing data by inspecting the missing data as lists.

>>>>

All measures can also be evaluated as a rolling estimate to detect changes in performance over time.
A rolling estimate is a sequence of estimates, each computed on a length-$w$-window of the data.
For the trending ratio, a rolling estimate with a backward-looking window is given by
\begin{equation*}
    \acc_t (\diffx, \diffy) \coloneqq \frac{\sum_{t^\star = t-w + 1}^{t} \ind{\diffxt[t^\star] \diffyt[t^\star] > 0}}{w} \quad t = w-1, \dots, T.\label{eq:acc_rolling}
\end{equation*}
Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the trending ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and~\ref{fig:trending_basic_4q_sample_color}.
Thus, the yearly course of the prediction's trending ratio can be detected.
The trending ability of the signal has a strong yearly, sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
This seasonal behavior cannot be distinguished in the two four-quadrant plots or in the trending ratio.

@@

In applications, signal data is often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
We refer to time steps, for which either signal or true values or both are unavailable, as missing values.
Data pairs with missing values can be excluded from the set $\mathcal{T}$ for the computation of the measures, and the measures are computed normally. 
However, systematical missing values could lead to a biased estimate of the trending ratio, for example, if signal data is omitted in times of high change and thus uncertainty.
In this case, the measures should be interpreted cautiously, and
inspection of missing data should accompany applications with missing data, for example, through a visual assessment.
In the applications in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}, only a few missing values occur, and we give information on the missing data by inspecting the missing data as lists.
